---
title: "Add topics to documents with topictagger"
author: Eliza Grames
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Add topics to documents with topictagger}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(topictagger)

```

# Introduction

# Tag all documents based on an ontology or topic scheme

## Read in documents

```{r}

```


## Construct scheme

```{r}

```

## Tag documents using standardized scheme

```{r}

```


# Tag new documents based on existing document tags

## Read in all documents

First, we read in the documents that have already been manually screened and had tags added to them.

```{r}


```

Next, we can read in new documents that need to be tagged.

```{r}

```

## Get document tags for new models

What's going on behind the scenes?

```{r}

```

# Tag all documents with no known topics

## Read in documents

```{r}

```


## Assign documents to topics

```{r}

```

## Explore topics

```{r}

```

